---
title: "Prediction-Assignment-Writeup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro 
The goal of this project is to predict the manner in which the subjects did the exercise. This is the "classe" variable in the training set.

The "classe" are:
- exactly according to the specification (Class A)
- throwing the elbows to the front (Class B)
- lifting the dumbbell only halfway (Class C)
- lowering the dumbbell only halfway (Class D)
- throwing the hips to the front (Class E)

## Load Libraries
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```

## Download Files
```{r}
url_train = 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
url_test = 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

download.file(url_train, 'train.csv')
download.file(url_test, 'test.csv')
```

## Load Files
```{r}
train_set = read.csv('train.csv', na.strings=c("NA","#DIV/0!", ""))
test_set = read.csv('test.csv', na.strings=c("NA","#DIV/0!", ""))
```

## Set seed
```{r}
set.seed(1909)
```

## Data Cleanup, convert classe to factor and Subset
```{r}
train_set <- train_set[,colSums(is.na(train_set)) == 0]
test_set <- test_set[,colSums(is.na(test_set)) == 0]

train_set$classe <- as.factor(train_set$classe)

train_set <- train_set[,-c(1:7)]
test_set <- test_set[,-c(1:7)]
```

## Splitting Data for cross-validation
```{r}
samples_cv <- createDataPartition(y=train_set$classe, p=0.75, list=FALSE)
train_cv <- train_set[samples_cv, ] 
test_cv <- train_set[-samples_cv, ]
```

## Prediction models

## random forest 
```{r}
model_rf <- randomForest(classe ~ ., data=train_cv, method="class")
predict_rf <- predict(model_rf, test_cv, type = "class")
```

```{r}
confusionMatrix(predict_rf, test_cv$classe)
```

### Decision tree
```{r}
model_dt <- rpart(classe ~ ., data=train_cv, method="class")
predict_dt <- predict(model_dt, test_cv, type = "class")
```

```{r}
confusionMatrix(predict_dt, test_cv$classe)
```

## Summary

The comparison of the confusion matrices shows that the RandomForest model performs better than the DecisionTree model and should therefore be selected.

The RandmForest model has an accuracy of 0.9949 with a 95% confidence interval of (0.9925, 0.9967). The DecisionTree model has an accuracy of 0.742 with a 95% confidence interval of (0.7296, 0.7542).

## Expected out-of-sample error
The expected out-of-sample error is estimated at 0.005, or 0.5%. The expected out-of-sample error is calculated as 1 - accuracy for predictions made against the cross-validation set. Our Test data set comprises 20 cases. With an accuracy above 99% on our cross-validation data, we can expect that very few, or none, of the test samples will be missclassified.

## Submission

```{r}
predict_sub <- predict(model_rf, test_set, type="class")
predict_sub

write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
write_files(predict_sub)
```